### 2024-02-12 starter_folder-main 4/paper/paper.qmd
updated intro and added 1 model

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

# Introduction

In recent years, the escalating cost of college tuition in the United States has emerged as a critical issue, intertwining with the broader narrative of rising income inequality. The paper by  Zhifeng Cai and Jonathan Heathcote addresses this pressing concern by examining the relationship between these two phenomena. Utilizing a competitive model of the college market, the study not only uncovers the mechanisms that link rising income inequality to increasing tuition costs but also sheds light on the consequent impacts on college attendance and social mobility.

Central to the paper's findings is the assertion that the surge in U.S. income inequality since the 1990s can account for more than half of the observed increase in average net tuition over the same period. This relationship underscores a critical feedback loop where income disparities not only affect individual capacity to afford higher education but also drive institutional behaviors around tuition setting. Moreover, the paper highlights the detrimental effects of rising tuition on college attendance rates, particularly among lower-income segments, positing significant implications for social mobility and equity.

To build upon the original work of Cai and Heathcote, our project will undertake a reproduction of the study, focusing on two or three of its core aspects. This will involve:

Recreating the Model: We will start by reproducing the competitive college market model used in the original study, ensuring our work is entirely reproducible. This step includes verifying the model's assumptions, computational methods, and data inputs.

# Data {#sec-data}

## Source

## Methodology

## Features

## Results

# Model

### Model justification

# Results

```{r}
#| echo: false
#| warning: false
#| message: false
#| 
# Load necessary libraries
library(ggplot2)
library(readxl)
library(dplyr)
library(tidyr)

# Set the file path
file_path <- "/Users/michaelfang/Downloads/Figure_1_data.xlsx"

# Read the data from the Excel file
data <- read_excel(file_path, sheet = "FIGURE 1 in PAPER", range = cell_cols("D:H"))

# Set column names manually
colnames(data) <- c('Academic Year', 'Public Sticker', 'Public Net', 'Private Sticker', 'Private Net')

# Check the structure of the data to make sure it's what we expect
str(data)

# Convert data types if necessary - all price columns should be numeric
data <- mutate(data, 
               `Public Sticker` = as.numeric(`Public Sticker`),
               `Public Net` = as.numeric(`Public Net`),
               `Private Sticker` = as.numeric(`Private Sticker`),
               `Private Net` = as.numeric(`Private Net`))

# Remove rows with NA values in any of the price columns
data <- na.omit(data)

# Ensure 'Academic Year' is a factor to maintain order in the plot
data$`Academic Year` <- factor(data$`Academic Year`, levels = unique(data$`Academic Year`))

# Reshape data to long format for ggplot2
data_long <- gather(data, key = "Price Type", value = "Amount", -`Academic Year`)

# Plot using the reshaped data
p <- ggplot(data_long, aes(x=`Academic Year`, y=Amount, group=`Price Type`, colour=`Price Type`)) +
  geom_line() +
  labs(title='Public vs. Private Sticker and Net Prices Over Time',
       x='Academic Year', y='Price in $') +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_colour_manual(values = c("Public Sticker"="blue", "Public Net"="green", "Private Sticker"="red", "Private Net"="purple"))

# Print the plot to see if it appears in the R console
print(p)

# Save the plot
ggsave("figure_1_plot.png", plot = p, width = 12, height = 8)
```

# Discussion

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

## Weaknesses and next steps

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

\newpage

# References